---
title: "Maternity Leave in NYC"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    source_code: embed
    runtime: shiny
---

```{r, include=FALSE}
library(tidyverse)
library(plotly)
library(flexdashboard)
library(shiny)
library(rsconnect)
library (tigris)
library(rgdal)
library(dplyr)
library(leaflet)
library(sp)


# rsconnect::deployApp("./shiny.Rmd")

# rsconnect::showLogs(appName="https://meb2308.shinyapps.io",streaming=TRUE)

#read in data frame for leave plots
charts_df = 
  read_csv("./data/merged_wfls.csv") %>% 
  drop_na(borough)

#create data frame with NAs for key socio-demographic variables removed
charts_df_2 = 
    read_csv("./data/merged_wfls.csv") %>% 
    drop_na(borough, race, job_type, leave_type, education, partner, family_income)

#load NYC map polygon files from map folder
zips = readOGR("./map/nyc_only_zips.shp", layer = "nyc_only_zips")

# develop colors palates for leaflet maps

leavewkspal <- colorNumeric(
  palette = "viridis",
  domain = as.numeric(zips$leaveweeks), na.color = NA)

wksunpaidpal <- colorNumeric(
  palette = "viridis",
  domain = as.numeric(zips$wksunpaid), na.color = NA)

lvtypepal <- colorFactor(topo.colors(3), zips$leavetype, na.color = NA)

parentpal <- colorFactor(topo.colors(3), zips$parenttype, na.color = NA)

edpal <- colorFactor(topo.colors(3), zips$edtype, na.color = NA)

racepal <- colorFactor(topo.colors(3), zips$race, na.color = NA)

jobtypepal <- colorFactor(topo.colors(3), zips$jobtypefix, na.color = NA)

neighborhoodpal  <- colorFactor(topo.colors(3), zips$X.neighborh, na.color = NA)

```

By Neighborhood
=====================================

Column {.sidebar}
-----------------------------------------------------------------------

```{r}
#choose a borough options

borough_choice = charts_df %>% 
  pull(borough) %>% unique() 

selectInput(
    "borough_choice",
    h3("Choose a borough"),
    choices = borough_choice)

```

Column {.tabset}
-----------------------------------------------------------------------
### Weeks of Paid and Unpaid Maternity Leave by Borough

```{r}
renderText({
  input[["borough_choice"]]
})

renderPlotly({
  
filtered_df <- 
  charts_df %>% 
  filter(
    borough == input$borough_choice
    ) %>% 
  group_by(leave_type) %>% 
  mutate(
    leave_type = factor(leave_type, levels=c("Paid", "Unpaid", "Both")),
    ) %>% 
  plot_ly(
    x = ~leave_type, y = ~leave_weeks, color = ~neighborhood,
    type = "box", colors = "viridis") %>%
  layout(
    boxmode = "group",
    title = "Weeks of Maternity Leave",
    xaxis = list(title = "Type of Leave"),
    yaxis = list(title = ""))
  
})

```

### Description

These boxplots display the distribution of weeks of maternity leave for women in NYC boroughs who gave birth in 2014.

"Paid" includes women who reported that all of the weeks of leave they took after giving birth were paid. "Unpaid" includes women who reported that all of the weeks of leave they took after giving birth were paid. "Both" includes women who took a mixture of paid and unpaid weeks of leave.

It is important to ....

===================================== 
By Key Socio-Demographics
=====================================

Column {.sidebar}
-----------------------------------------------------------------------

```{r}
#pull in relevant data

race = charts_df_2 %>% 
  pull(race) %>% unique()

education = charts_df_2 %>% 
  pull(education) %>% unique()

partner = charts_df_2 %>% 
  pull(partner) %>% unique()

employment = charts_df_2 %>% 
  pull(job_type) %>% unique()

family_income = charts_df_2 %>% 
  pull(family_income) %>% unique()

#Choose two options to plot in bar chart format

selectInput(
  "choice_1",
  h3("Choose x-axis variable"),
  choices = c("Race", "Education", "Co-parenting Status", "Employment Type", "Family Income"),
  selected = "Race")

conditionalPanel(
  condition = "input.choice_1 == 'Race'",
  checkboxGroupInput(
  "race",
  h3("Race"),
  choices = race))

conditionalPanel(
  condition = "input.choice_1 == 'Employment Type'",
  checkboxGroupInput(
  "employment",
  h3("Employment Type"), 
  choices = employment))

conditionalPanel(
  condition = "input.choice_1 == 'Co-parenting Status'",
  checkboxGroupInput(
  "partner",
  h3("Co-parenting Status"),
  choices = partner))

conditionalPanel(
  condition = "input.choice_1 == 'Education'",
  checkboxGroupInput(
  "education",
  h3("Education"),
  choices = education))

conditionalPanel(
  condition = "input.choice_1 == 'Family Income'",
  checkboxGroupInput(
  "family_income",
  h3("Family Income"), 
  choices = family_income))

selectInput(
  "choice_2",
  h3("Choose color group variable"),
  choices = c("Race", "Education", "Co-parenting Status", "Employment Type", "Family Income"),
  selected = "Family Income")

```

Column {.tabset}
-----------------------------------------------------------------------

### Weeks of Maternity leave by Sociodemographic Characteristics

```{r}
renderPlotly({
  
  plots <-
  charts_df %>%
  group_by(input$choice_1, input$choice_2) %>% 
  plot_ly(
    x = ~input$choice_1, y = ~leave_weeks, color = ~input$choice_2,
    type = "box", colors = "viridis") %>% 
  layout(
    title = paste("Weeks of Maternity Leave by", input$choice_1, "and", input$choice_2),
    xaxis = list(title = input$choice_1),
    yaxis = list(title = ""))
})
```

### Description

This plot...

===================================== 
NYC Maps
===================================== 

Column {.sidebar}
-----------------------------------------------------------------------

```{r}
# Choose map options

borough_choice = charts_df %>% 
  pull(borough) %>% unique()
  
race = charts_df_2 %>% 
  pull(race) %>% unique()

education = charts_df_2 %>% 
  pull(education) %>% unique()

partner = charts_df_2 %>% 
  pull(partner) %>% unique()

employment = charts_df_2 %>% 
  pull(job_type) %>% unique()

family_income = charts_df_2 %>% 
  pull(family_income) %>% unique()

# Choose one option for map in addition to Borough

selectInput(
  "geography",
  h3("Choose a geography"),
  choices = list("All of NYC", "Result by borough"))

conditionalPanel(
  condition = "input.geography != 'All of NYC'",
    selectInput(
    "borough_choice",
    h3("Choose a borough"),
    choices = borough_choice))

selectInput(
  "choice_1",
  h3("Choose another variable to map"),
  choices = list("Race", "Education", "Co-parenting Status", "Employment Type",  "Family Income"))

```

Column
----------------------------------------------------------------------

### Maps 
```{r}
renderLeaflet({
  
wksunpaid = 
  leaflet(zips) %>% 
    addPolygons(color = "#444444", weight = 1, smoothFactor = 0.5,
    highlightOptions = highlightOptions(color = "white", weight = 2,
      bringToFront = TRUE),
    opacity = 1.0, fillOpacity = 0.5,
    fillColor = ~wksunpaidpal(as.numeric(zips$wksunpaid))) %>% 
  addProviderTiles(providers$CartoDB.Positron)

})

renderLeaflet({
leaveweeks = 
leaflet(zips) %>% 
    addPolygons(color = "#444444", weight = 1, smoothFactor = 0.5,
    highlightOptions = highlightOptions(color = "white", weight = 2,
      bringToFront = TRUE),
    opacity = 1.0, fillOpacity = 0.5,
    fillColor = ~leavewkspal(as.numeric(zips$leaveweeks))) %>% 
  addProviderTiles(providers$CartoDB.Positron)

})
```

Column
----------------------------------------------------------------------

```{r}
renderText({
  input[["borough_choice"]]
  input[["choice_1"]]
})

renderLeaflet({
  race = 
    leaflet(zips) %>% 
    addPolygons(color = "#444444", weight = 1, smoothFactor = 0.5,
    highlightOptions = highlightOptions(color = "white", weight = 2,
      bringToFront = TRUE),
    opacity = 1.0, fillOpacity = 0.5,
    fillColor = ~racepal(zips$race)) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
    addLegend("bottomright", pal = racepal, values = ~race,
    title = "Race",
    opacity = 0.5)
  
})

renderLeaflet({
  edtype = 
  leaflet(zips) %>% 
    addPolygons(color = "#444444", weight = 1, smoothFactor = 0.5,
    highlightOptions = highlightOptions(color = "white", weight = 2,
      bringToFront = TRUE),
    opacity = 1.0, fillOpacity = 0.5,
    fillColor = ~edpal(zips$edtype)) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
     addLegend("bottomright", pal = edpal, values = ~edtype,
    title = "Education Level",
    opacity = 0.5)
  
})

renderLeaflet({
  parenttype = 
    leaflet(zips) %>% 
    addPolygons(color = "#444444", weight = 1, smoothFactor = 0.5,
    highlightOptions = highlightOptions(color = "white", weight = 2,
      bringToFront = TRUE),
    opacity = 1.0, fillOpacity = 0.5,
    fillColor = ~parentpal(zips$parenttype)) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
    addLegend("bottomright", pal = parentpal, values = ~parenttype,
    title = "Co-parenting Status",
    opacity = 0.5)
})

renderLeaflet({
  jobtype = 
    leaflet(zips) %>% 
    addPolygons(color = "#444444", weight = 1, smoothFactor = 0.5,
    highlightOptions = highlightOptions(color = "white", weight = 2,
      bringToFront = TRUE),
    opacity = 1.0, fillOpacity = 0.5,
    fillColor = ~jobtypepal(zips$jobtypefix)) %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
    addLegend("bottomright", pal = jobtypepal, values = ~jobtypefix,
    title = "Employment Type",
    opacity = 0.5)
  
})
```


